Spatial Analysis

### 2021-10-26 Scripts/spatial_analisis.Rmd
Set reference points for each stock

# Load all sp in one table
historic_spp <- bind_rows(lapply(taxon_sw, fread)) %>% 
  left_join(grids,
            by = c("index","lon","lat")
            ) %>%
  filter(!is.na(spatial)) %>% 
  mutate(
    season = ifelse(str_detect(region,"Spring"),"Spring","Fall")
  ) %>% 
  filter(spp == "Centropristis striata")

@@

# Interpolation data
interpol_result <- historic_spp %>% 
  mutate(
    label = ifelse(year >= 1980 & year <= 2001,"Reference",
                   ifelse(year > 2001,"Today",NA)
    )
  ) %>% 
  filter(!is.na(label)) %>% 
  group_by(label,index,spp,season,lat,lon) %>% 
  summarise(mean_cpue = mean(value,na.rm=T)) %>% 
  filter(!is.na(mean_cpue))

@@

# Estimate total interpolation
total_inter <- historic_spp %>% 
  mutate(
    label = ifelse(year >= 1980 & year <= 2001,"Reference",
                   ifelse(year > 2001,"Today",NA)
    )
  ) %>% 
  filter(!is.na(label)) %>% 
  group_by(spp,index,season) %>% 
  summarise(mean_cpue = mean(value,na.rm=T)) %>% 
  group_by(spp,season) %>% 
  summarise(total_cpue = sum(mean_cpue, na.rm = T))

# Estimate centroid based on top 10 grids with higer abundance
centroid_db <- historic_spp %>% 
  mutate(
    label = ifelse(year >= 1980 & year <= 2001,"Reference",
                   ifelse(year > 2001,"Today",NA)
    )
  ) %>% 
  filter(!is.na(label)) %>% 
  group_by(label,index,spp,season,lat,lon) %>% 
  summarise(mean_cpue = mean(value,na.rm=T)) %>% 
  filter(!is.na(mean_cpue)) %>% 
  left_join(total_inter) %>% 
  mutate(
    per_cpue = round((mean_cpue/total_cpue)*100,2)
  ) %>% 
  group_by(spp,label,season) %>% 
  top_n(10,per_cpue) %>% 
  group_by(spp,label,season) %>% 
  summarise(
    lat = mean(lat),
    lon = mean(lon)
  )

@@

inter_map <- ggplot() +
  geom_tile(data = subset(interpol_result, mean_cpue == 0),
            aes(
              x = lon,
              y = lat
            ),
            fill = "grey90",
            color = "grey90"
  ) + 
  geom_tile(data = subset(interpol_result, mean_cpue > 0),
            aes(
              x = lon,
              y = lat,
              # fill = log10(mean_cpue),
              # color = log10(mean_cpue)
              fill = mean_cpue,
              color = mean_cpue
            )
  ) + 
  geom_point(data = centroid_db,
             aes(
               x = lon,
               y = lat
             ),
             shape = 3,
             color = "white"
  ) + 
  geom_sf(data = land_sf,aes()) +
  labs(x = "Longitude",
       y = "Latitude") +
  # ggsflabel::geom_sf_label_repel(data = land_sf, aes(label= abrev), 
  #                                  size = 4,
  #                                  box.padding = 0.10,
  #                                  hjust = 1) +
  # Viridis option
scale_fill_viridis("CPUE",
                   limits = c(0,40),
                   # limits = c(-5,1), # for log10
                   direction = -1,
                    end = 0.9
                   ) +
scale_color_viridis("CPUE",
                    limits = c(0,40),
                    # limits = c(-5,1), # for log10
                    direction = -1,
                    end = 0.9
                    ) +
# wes anderson option
# scale_fill_gradientn("CPUE",
#                      colours = pal,
#                      limits = c(0,40) # For cpue
#                      # limits = c(-5,1) # for log10
#                      
# ) +
#   scale_color_gradientn("CPUE",
#                         colours = pal,
#                         limits = c(0,40) # for cpue
#                         # limits = c(-5,1) # for log10
#   ) +
  facet_grid(spp~season+label) +
  my_ggtheme_m(map_type = "na", leg_pos = "right",ax_tx_s = 7,ax_tl_s = 10,leg_tx_s = 12) #+
# theme(legend.key.width = unit(3,"line"))

ggsave("interpolation_vi.jpg",
       inter_map,
       path = my_path("R","Figures"),
       width = 9,
       height = 7)

@@

state_fit <- historic_spp %>% 
  group_by(state,year,season,spp,spatial) %>% 
  summarise(state_value = sum(value,na.rm= T), .groups = "drop") %>% 
  # View()
  left_join(total_fited,
            by = c("year","season","spp","spatial")) %>%
  mutate(percentage = state_value/total_value*100,
         label = ifelse(year >= 1980 & year <= 2001,"Reference",
                         ifelse(year > 2001,"Today",NA)
                         )
         ) %>% 
  group_by(state,label,season,spp,spatial) %>% 
  summarise(mean_per = round(mean(percentage)),.groups = "drop") %>% 
  #Only show results for spring
  filter(!is.na(label))

@@

# group by state
state_fit <- historic_spp %>% 
  group_by(state,year,region,spp,spatial) %>% 
  summarise(state_value = sum(value,na.rm= T), .groups = "drop") %>% 
  left_join(total_fited,
            by = c("year","region","spp","spatial")) %>%
  mutate(percentage = state_value/total_value*100) %>% 
  group_by(state,region,spp,spatial) %>% 
  mutate(RMean = zoo::rollmean(x = percentage, 
                            10,
                            fill = NA,
                            na.rm=T)
    ) %>% 
  #Only show results for spring
  filter(str_detect(region,"Spring"),
         !is.na(RMean) # Remove NAs from rollmean
         ) %>%  
  left_join(state_lat) %>% 
  mutate(
         spatial = ifelse(spatial == "fp","Fishing ports",
                          ifelse(spatial == "sw","State waters","Difference")
         )
  )

# Plot me
ssp_area <- ggplot(state_fit) +
  geom_area(
    aes(
      x = year,
      y = round(RMean),
      fill = order
      # fill = state
    )
  ) +
  annotate(geom= "rect",
           xmin = 1980, xmax = 2001, 
           ymin = 0, ymax = 100,
            fill = "gray90",
           colour = NA,
           alpha = 0.5) +
  ylab("Propotion (10 yrs running average)") +
  # viridis::scale_fill_viridis(discrete = T, alpha = 0.8) +
  scale_fill_manual(
    "State\n(Latitudinal order)",
    values = state_pallet,
    labels = state_fit %>% arrange(order) %>%  pull(state) %>% unique()
  ) +
  facet_grid(spp~spatial) +
  scale_x_continuous(breaks = seq(1979,2015,9),limits = c(1979,2015))+
  xlab("Year") +
  my_ggtheme_p(leg_pos = "right", hjust = 0.9,ax_tx_s = 14,leg_tx_s = 16, leg_tl_s = 18,ax_tl_s = 18,facet_tx_s = 16) +
  theme(legend.key.width = unit(1,"line"),
        legend.title.align=0.5)

>>>>

# Load all sp in one table
historic_spp <- bind_rows(lapply(taxon_sw, fread)) %>% 
  left_join(grids,
            by = c("index","lon","lat")
  ) %>%
  filter(!is.na(spatial)) %>% 
  mutate(
    season = ifelse(str_detect(region,"Spring"),"Spring","Fall"),
    label = ifelse(spp == "Paralichthys dentatus" & year >= 1980 & year <= 1986,"Reference",
                   ifelse(spp == "Stenotomus chrysops" & year >= 1988 & year <= 1992,"Reference",
                          ifelse( spp =="Centropristis striata" & year >= 1980 & year <= 2001,"Reference",
                                 ifelse(year > 2001,"Today",NA)
                          )
                   )
    )
  )

@@

# Interpolation data
interpol_result <- historic_spp %>% 
  filter(!is.na(label)) %>% 
  group_by(label,index,spp,season,lat,lon) %>% 
  summarise(mean_cpue = mean(value,na.rm=T)) %>% 
  filter(!is.na(mean_cpue))

@@

# Estimate total interpolation
total_inter <- historic_spp %>% 
  filter(!is.na(label)) %>% 
  group_by(spp,index,season) %>% 
  summarise(mean_cpue = mean(value,na.rm=T)) %>% 
  group_by(spp,season) %>% 
  summarise(total_cpue = sum(mean_cpue, na.rm = T))

# Estimate centroid based on top 10 grids with higer abundance
centroid_db <- historic_spp %>% 
  filter(!is.na(label)) %>% 
  group_by(label,index,spp,season,lat,lon) %>% 
  summarise(mean_cpue = mean(value,na.rm=T)) %>% 
  filter(!is.na(mean_cpue)) %>% 
  left_join(total_inter) %>% 
  mutate(
    per_cpue = round((mean_cpue/total_cpue)*100,2)
  ) %>% 
  group_by(spp,label,season) %>% 
  top_n(10,per_cpue) %>% 
  group_by(spp,label,season) %>% 
  summarise(
    lat = mean(lat),
    lon = mean(lon)
  )

@@

inter_map <- ggplot() +
  geom_tile(data = subset(interpol_result, mean_cpue == 0),
            aes(
              x = lon,
              y = lat
            ),
            fill = "grey90",
            color = "grey90"
  ) + 
  geom_tile(data = subset(interpol_result, mean_cpue > 0),
            aes(
              x = lon,
              y = lat,
              fill = log10(mean_cpue),
              color = log10(mean_cpue)
              # fill = mean_cpue,
              # color = mean_cpue
            )
  ) + 
  geom_point(data = centroid_db,
             aes(
               x = lon,
               y = lat
             ),
             shape = 3,
             color = "white"
  ) + 
  geom_sf(data = land_sf,aes()) +
  labs(x = "Longitude",
       y = "Latitude") +
  # ggsflabel::geom_sf_label_repel(data = land_sf, aes(label= abrev), 
  #                                  size = 4,
  #                                  box.padding = 0.10,
  #                                  hjust = 1) +
  # Viridis option
# scale_fill_viridis("CPUE",
#                    limits = c(0,40),
#                    # limits = c(-5,1), # for log10
#                    direction = -1,
#                     end = 0.9
#                    ) +
# scale_color_viridis("CPUE",
#                     limits = c(0,40),
#                     # limits = c(-5,1), # for log10
#                     direction = -1,
#                     end = 0.9
#                     ) +
# wes anderson option
scale_fill_gradientn("CPUE",
                     colours = pal,
                     # limits = c(0,40) # For cpue
                     limits = c(-5,1) # for log10

) +
  scale_color_gradientn("CPUE",
                        colours = pal,
                        # limits = c(0,40) # for cpue
                        limits = c(-5,1) # for log10
  ) +
  facet_grid(spp~season+label) +
  my_ggtheme_m(map_type = "na", leg_pos = "right",ax_tx_s = 7,ax_tl_s = 10,leg_tx_s = 12) #+
# theme(legend.key.width = unit(3,"line"))

ggsave("interpolation_wa.jpg",
       inter_map,
       path = my_path("R","Figures"),
       width = 9,
       height = 7)

@@

state_fit <- historic_spp %>% 
  group_by(state,year,label,season,spp,spatial) %>% 
  summarise(state_value = sum(value,na.rm= T), .groups = "drop") %>% 
  # View()
  left_join(total_fited,
            by = c("year","season","spp","spatial")) %>%
  mutate(percentage = state_value/total_value*100) %>% 
  group_by(state,label,season,spp,spatial) %>% 
  summarise(mean_per = round(mean(percentage)),.groups = "drop") %>% 
  #Only show results for spring
  filter(!is.na(label))

@@

# group by state
state_fit <- historic_spp %>% 
  group_by(state,year,label,region,spp,spatial) %>% 
  summarise(state_value = sum(value,na.rm= T), .groups = "drop") %>% 
  left_join(total_fited,
            by = c("year","region","spp","spatial")) %>%
  mutate(percentage = state_value/total_value*100) %>% 
  group_by(state,region,spp,spatial) %>% 
  mutate(RMean = zoo::rollmean(x = percentage, 
                            10,
                            fill = NA,
                            na.rm=T)
    ) %>% 
  #Only show results for spring
  filter(str_detect(region,"Spring"),
         !is.na(RMean) # Remove NAs from rollmean
         ) %>%  
  left_join(state_lat) %>% 
  mutate(
         spatial = ifelse(spatial == "fp","Fishing ports",
                          ifelse(spatial == "sw","State waters","Difference")
         )
  )

# Shaded period 
historic_periods <- historic_spp %>% 
  filter(label == "Reference") %>% 
  group_by(spp,label,year) %>% 
  summarise(n = n()-n()+100) %>% 
  mutate(period = "Reference period")

# Plot me
ssp_area <-
  ggplot(state_fit) +
  geom_area(
    aes(
      x = year,
      y = round(RMean),
      fill = order
      # fill = state
    )
  ) +
  geom_area(data = historic_periods,
             aes(
               x = year,
               y = n,
               color = period
             ),
             # fill = "gray90",
             alpha = 0.3,
            stat = "identity"
             ) +
  # annotate(geom= "rect",
  #          xmin = 1980, xmax = 2001, 
  #          ymin = 0, ymax = 100,
  #           fill = "gray90",
  #          colour = NA,
  #          alpha = 0.5) +
  ylab("Propotion (10 yrs running average)") +
  # viridis::scale_fill_viridis(discrete = T, alpha = 0.8) +
  scale_fill_manual(
    "State\n(Latitudinal order)",
    values = state_pallet,
    labels = state_fit %>% arrange(order) %>%  pull(state) %>% unique()
  ) +
    scale_color_manual("",values = "grey90"
  ) +
  facet_grid(spp~spatial) +
  scale_x_continuous(breaks = seq(1979,2015,9),limits = c(1979,2015))+
  xlab("Year") +
  my_ggtheme_p(leg_pos = "right", hjust = 0.9,ax_tx_s = 14,leg_tx_s = 16, leg_tl_s = 18,ax_tl_s = 18,facet_tx_s = 16) +
  theme(legend.key.width = unit(1,"line"),
        legend.title.align=0.5)
